Corporate Security Solutions for BYOD.\\ A Novel User-Centric and Self-Adaptive System

### 2015-04-15 MUSES_vs_BYOD_solutions.tex
Revisado el inglés
He cambiado muchos focus in->focus on, for verbo->for verboing, typos y cosillas de ese estilo. También muchas comas y cambiadas algunas frases.
He puesto algunas dudas en comentarios, pero aún no lo he leido en profundidad. Mañana sigo, así que @JJ @unintendedbear @amorag podéis seguir poniéndome issues si queréis :)

The way on how data has been stored and accessed in the companies has completely changed along the years. The use of owned servers being accessed by desktop PCs (and maybe laptops) from inside the company facilities has been transformed into the distribution of these data among a number of machines, even not all belonging to the company, and working over the famous Cloud Computing environment. Besides, being stored in the cloud or not, data are being consulted and modified through a wide amount of devices, some of them owned by the company's users. This is the so-called Bring Your Own Device (BYOD) philosophy. It is becoming highly successful due to the impact that portable devices (such as smartphones and tablets) are having in the market.
Data security and privacy are key factors for a company, thus, to protect them, it is usual to define Organisational Security Policies. Their definition is nowadays a very difficult problem, since the BYOD tendency means that several factors must be considered \cite{Opp_Security11}, most of them previously ignored or non considered in security systems, for instance the current mixture between personal and professional information in these devices (the user could navigate inside social networks where there could be friends and also company partners or clients).

In this scenario several solutions have arisen in order to manage the corporate security. Some solutions are focused only in smartphones, while others are thought for laptops; also some are implemented for a certain platform, but others consider multiplatform. However most of them try to be non-intrusive (regarding the users' personal data), friendly and easy to use.

Moreover, it has been demonstrated that people are the main hazard regarding the company security \cite{Adams_Users05}, so in this situation, some monitorization and security-aimed applications are arising, aiming to cope with the concept of seamless working experience on different devices. 
This concept is a methodology of work which allows users to start/continue a working session over multiple devices and locations without any significant loss of data. This new situation has a big impact from the point of view of the security \cite{Schu_SecPatterns05}, since the company's data borders have changed in the last years so now the users can access significant data from outside the enterprise, and possibly through a non absolutely secure channel.

Thus, this work introduces a novel (still in development) system named MUSES, from Multiplatform Usable Endpoint Security, which is a device-independent end-to end user-centric tool, based in a set of security rules defined as specifications of the Corporate Security Policies, but with the ability of `learning' from the user's past behaviour and adapt, even inferring new ones, the set of rules in order to deal with potential future security incidents due to the user's. Then the system will react, in a non-intrusive way, to the potentially dangerous sequence of actions (events) that he or she is conducting at any time.

@@

On the other hand, employee-owned devices like smartphones, can be both used for personal matters at work, and for continue working outside of the workplace. As the name says, smartphones are more than simple mobile phones, and people who use them in their works have the possibility of maintaining a good balance between work and private life. For this reason, the risk of uncontrolled devices accessing to corporate assets in unsafe conditions, due to the number of new risks which are linked to smartphones \cite{gangula2013survey}, is bigger.

@@

Many tools for companies, as well as for devices, which have adopted the BYOD have been released in the past four years. This way, and
more focused on the enterprise, there are some tools which offer the CSO ways to control the devices which enter in the system, for instance, and also to protect the employees' data by means of data encryption and data protection by requiring the use of strong and secure passwords. Other tools for managing a BYOD situation add to their features guidelines for the CSOs to develop good ISPs. On the other hand, many solutions have been presented which are more focused on the device side, although they implement also the server side. Some of those tools have influenced the development of the MUSES system  itself (Section \ref{sec:muses}), which can complement them, as adds other features that go beyond the state of the art. The present section introduces the most relevant features of these products, as they can be considered related to the MUSES objectives.

@@

One of the first companies who supported the BYOD model was IBM \cite{IBM_tool}, as they recognized the increase of employees who brought their personal smartphones or tablets into the workplace \cite{ibm11}. IBM has different solutions divided into ``technology solutions'', and ``services'', and almost all are are mainly focused in the management of the devices in the system. Then, it might seem clear that the first disadvantage is not including all features in one system, but to make the companies to choose between one or another, or invest even more resources instead.

Having a look into what IBM offer as ``services'', three services can be found:

\begin{itemize} 
	\item \textit{Mobile Enterprise Services}: Is offered as an integrated solution for smartphones, and tablets as well, but then it is divided in different subservices that the company has to acquire separately. Some of them are related to cloud computing, such as events and log management as a cloud security service, cloud backup, or the so-called `hosted vulnerability management'. The last one is specially interesting for the scope of this paper, given the fact that it seems to implement one of the most important advantages of the MUSES system, the self-adaptation. IBM, by performing deeper scans over the security incidents data, either if they were successful (computer forensics analysis) or not (the device was enough secure), they claim to be able to recognise new vulnerabilities - or threats - with enough accuracy. The main difference with MUSES in this case, is that MUSES includes this feature in the system, without the need of purchasing the service separately, and configuring the different parts. Also, the fact that IBM software is not open source as MUSES system is, to verify the success of this feature is highly difficult, as they do not present statistics in their website either.
	\item \textit{Hosted Mobile Device Security Management}: A particular service of the Mobile Enterprise Services. It includes itself a number of subservices, going from tools to make easy to develop secure mobile apps, to an interface for managing the devices in the system, checking their configuration security level. Again, what is offered by IBM as `services', are included as default features in the MUSES system.
	\item \textit{Enterprise Wireless Networks}: Devoted to provide secure channels for communications. This is related with ensuring employees that any connection they might perform to access company assets, is properly secured, at the time it delivers the expected performance. Thus, this service is, in some sense, adaptive to the situation, yet only in terms of a secure connection and not evolving the security policies, like MUSES. 
\end{itemize}

@@

Contrarily to IBM's solution services, Sophos offer a whole product for securing BYOD environments, which is called Mobile Control \cite{Sophos_tool}. It is oriented to IT administration for mobile devices, and it can be deployed in two ways: on-premises, which means that all data and services remain in the servers of the company, and as Software as a Service (SaaS), so that the services are provided through Lightweight Directory Access Control (LDAP). LDAP is an application protocol for accessing and maintaining distributed directory information services over an Internet Protocol (IP) network.

The tool manages all employees' smartphones and tablets from a single-based console. This console monitors the devices since initial set up and registration, until they log off. Other features are similar to IBM's product, adding some new such as the incorporation of a service called Malware and Web protection, so that there is no need for having a previously installed antivirus in the devices.

With regard to compliance enforcement, the main goal of this Mobile Control tool is to give more importance to company security, instead of giving more flexibility to the users. But it has been demonstrated by Herath and Rao in \cite{SecPolPenalty09} that this practise can result in bad (even not on purpose) users' behaviour. Thus, company BYOD initiative should include an acceptable use policy to ensure that the users are aware of the measures that the company would take if a device breaches any ISPs. Sophos aims to reach this by doing three main tasks: First, by allowing setting up user and group-based security policies separately. Secondly, by a risk mitigation in which the actions to perform can be set according to the severity of the breach. For minor cases, the company may want to simply inform to the user, but if sensitive data is at risk, a remote wipe may be the chosen option. Finally, compliance check, by constantly performing scans to detect malware and putting the devices in quarantine when they are found infected.

Thus, given the fact that this tool is more focused in the company than in the user, it is worth to compare it to the MUSES system, as the latter tries to focus in the users, being user-centric. This means that is preferred to let the user know about a risky situation, rather than making them feel ``watched''.

@@

The main feature of KNOX is the use of different containers, or environments, for business and personal sides. Each one even includes its own graphic configuration (wallpapers, colours, and so on), in order to be more easily recognised and distinguished by the user.
A password is required in order to enter into the business side and, once \textit{logged in} in this container, no more passwords are needed for the business applications. The applications approved by the company IT department must meet Samsung security standards and, though it has been said that KNOX is a device-focused solution, it offered an API. This API would allow the company to access to over almost 205 predefined IT policies. With regard to information protection methods, data files saved by applications of each environment are encrypted with AES 256-bit algorithm, in such manner that only the appropriate container can access these files. In the same way, the user is not allowed to share data between the two environments.

Features in Blackberry Balance are similar. This security package was announced as a feature of BlackBerry 10. Nevertheless, it is available with BlackBerry Enterprise Service 10, which is a device management, security and app management for BlackBerry, iOS and Android devices. It is necessary to activate BlackBerry Balance for having available some security features, all related or similar to the aforementioned. For instance, a message is displayed when the user tries to copy work data and then paste it into personal apps. Also, user attempting for actions that are not permitted in the company ISP, or may cause secure work information to be in contact with personal applications, these actions won't be permitted.

@@

This system will be strongly related with the Google Play market, as the applications will be `categorized'. Thus, Android Work will provide the CSOs with a tool to define which apps would be allowed for being installed by the users and considered as corporate-related applications. Several enterprise-security aimed applications will be offered in the market, so the IT would decide which ones will be installed (even as a bundle) in the employees' devices, and also, they can manage the updates of these apps, in order to ensure that all the employees are up to date. For instance, one app could manage the creation of personal/work profiles, so the user just could access corporate assets after login in into the app. Moreover, the enterprise IT department could define specific security policies on these apps, in addition to the inherent protection that the container provides. However, the specifications of this tool \cite{AndroidWork_tool} do not include the ``self-adaptation'' as a feature, as the MUSES system does.

Even though it does not analyse the system information for security rules evolution purposes, it offers policy management. Thus, the system will provide a framework for IT staff to manage business devices but also, and more importantly, personal devices being used in a BYOD context (The CSO would give the employee an activation code in order to `connect' the smartphone to the enterprise and use this Android Work).
Therefore, IT admins will be able to specify which Google Play apps will be available for users to install through this work profile, including being able to provision apps to specific individuals or groups (inside the company).

@@

\begin{description}
  \item[Mobile Device Management] that is used to manage both user and corporate owned devices, providing support for Android and iOS at the moment. It should be noted that this tool allowes the tracking of every enrolled device, as well as obtaining reports and analytics of their use.
  \item[Mobile Application Management] With regard to the software, this platform is able to allow or deny the use of applications on enrolled devices based on the role of the user or policies, thus restricting the use of some apps to certain users.
  \item[Enterprise App Store] This store provides users with both enterprise and public applications approved by the company.
  \item[Mobile Data Security] WSO2 also allows the user's data to be encrypted via password.
\end{description}

@@

\begin{itemize}
\item It offers a centrally deployment, administration and securing of all mobile devices regardless of their OS.
\item It manages and enforces security policies `over the air' for both corporate and employee-owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe/selective wipe corporate data email/contacts on the device if lost or stolen.
\end{itemize}

@@

It is also important to point out that Citrix differenciates between apps used by the user privately and those used for business, locating both on a secure mobile container that is encrypted, and can be locked remotely for safety reasons. Another feature of Citrix is that it uses dedicated micro VPN to connect to Citrix-protected backend services.

Once the main tools and systems in this area have been introduced, in the following section our own system (still in development) is presented, giving an overview on its general architecture. Moreover we devote a subsection to describe its main features: the use of Data Mining + Machine Learning techniques, and also its self-adaptation ability using Computational Intelligence methods, which compose the real difference with all these tools.

@@

As a summary, this application includes two modules: a \textit{controller} and an \textit{actuator}. The first one monitors the environment (context) and the user's behaviour, translating his/her actions into a sequence of events. These events, along with the patterns defining the user's conduct, are processed by the system in real-time by means of a Risk and Trust Analysis Engine (RT2AE) and an Event Correlation module. Then, a decision is taken in the corporate security operations centre (SOC) side, considering the RT2AE output and the set of security rules adapted to that specific user and context. The corresponding feedback is communicated to the user through the \textit{actuator}, which is also in charge of triggering the recommended actions to stop the user's or application's doings, in case it is required.

The designed MUSES architecture is shown in Figure \ref{fig:architecture}. It is a \textit{client/server} approach in which the \textit{client} program will be installed in every user's mobile or portable device, independently of the platform (operating system and type of device). The \textit{server} side would be installed in the corporate SOC. Both sides are connected through a secure channel (using HTTPS) over Internet. In that figure just the high-level components in each part are shown, along with the information flows labelled as 'Info XX'.

@@

The rationale for this approach is based on two main reasons: 1) there is need for a high computational power in order to perform the event correlation and self-adaptation processes, so a quite powerful machine should be used (server); 2) there are two clearly separated parts in the system, namely the users (client) and the enterprise (server).

The system considers two working modes for every device: online and offline. In the \textit{online} mode the device can connect with the MUSES server, so it can request the server to make a decision. On the other hand, in \textit{offline} mode the server cannot be reached by the device (there is not an available connection between them), so all the decisions should be made locally.
This will be done using an up-to-date copy on the Decision Table inside the Local database, which will contain decisions for any type of detected event, along with default actions to be performed if no rule is met. Anyway, in this mode, the gathered information by the sensors in the device will be stored for later submission to the server side (when a connection is available), in order to be processed in the knowledge refinement process.
Moreover, when switching from offline to online mode, the device will receive an updated copy of the Decision Rules (the so-called Device Policies), as a kind of antivirus database, in order to work again perfectly in offline mode (if the connection is lost).

All data in MUSES will be encrypted.

@@

The rest of the components are: the \textit{MUSES User Interface (UI)}, the application through which the user interacts with the system; and the \textit{Connection Manager} which controls the communications between client and server sides. 
In addition, there are two types of applications considered in this system: on the one hand the \textit{MUSES Aware App}, which is an application adapted to MUSES, so the system can directly interact with it (monitoring and acting). This application must be implemented using the MUSES API (Application Program Interface)\footnote{The MUSES API will be defined in the project, so for every application desired to be MUSES Aware, it should be implemented using this.}. On the other hand the \textit{Non MUSES Aware App} is that which MUSES cannot directly interact with. Usually it will be accessed through the operating system (OS). So the type of information to gather from them and also the type of actions to do over them will vary depending on the OS. In most of the cases MUSES could just detect/recognize very specific events, and also do very limited actions to prevent or block them (maybe showing a screen on the top until something is accomplished by the user). This is due to the usual forbiddance or limitation of interactions with other applications that the OSs have.

@@

There are some other components, namely: the \textit{Security Policies/Risk Management} tool, that lets the company's chief security officer (CSO) to define and manage Security Policies and Rules in a friendly way. It also lets the management of risk-related information, useful for the RT2AE process, such as the assets' values. The \textit{Privacy Enhancing System} which is a module aimed to fit with the legal compliance of the system regarding the user's data anonymisation and retaining of these data in the system. The \textit{User, Context, Event Data Receiver} is devoted to receive data from the device side (events, user-related, etc) and to distribute them among the components (storing in the database, and requesting the Event Processor to start the correlation task). Finally, there is another \textit{Connection Manager} which controls the communication with the device side.

@@

Another important fact is that MUSES will count with a human controller, normally the Chief Security Officer (CSO) of the company, who will supervise the system activity by means of logs or feedback messages. Thus, adapted and inferred security rules will not be, in principle, directly added to the current set of rules. Instead, they will be proposed as draft rules to this controller, in order that he/she will accept them if they are interesting and correct. It is planned that the system will be able to `learn' from this decisions so, after a so-called training or `warm up' period, the rules would be directly accepted (included in the set) or rejected autonomously by the MusKRS.

@@

The DM/ML techniques will process the so-called patterns, which in this context correspond to events (and their related information) produced by the users' interactions with the system. As stated, these events will be captured by the sensors in the users' devices and transmitted to the server to be processed and stored in the system database. The methods to be applied are:

@@

\item \textit{Classification} \cite{classification_67}: 
This technique tries to train a model (classifier) able to associate every pattern in the dataset to a class, so that the model could be used for assign a class for further incoming patterns with an unknown category.
For instance, it could look for events (patterns) that had been previously  marked as `allowed' or `denied' (according to the ISPs). When a new event arises, if it has not an assigned decision, the classifier should provide one based on the similarity with previous (and already labelled) patterns.

@@

\item \textit{Feature Selection} \cite{FeatureSelection_Guyon03}: 
It consists on extract the most important features/variables from the data. This could be useful if we want to discard non-key features, which could be interesting in order to reduce the database weight, for improving the performance of other techniques (such as classification or clustering), and even to improve the performance of whole the system, since less information would be gathered and transmitted.

@@

MUSES will use different EC approaches, initially, two in the DM/ML part of the process, and three in the rule-refinement/adaptation phase. They are based in Genetic Programming (GP) \cite{GP_Koza92}, and Genetic Algorithms (GAs) \cite{GAs_Goldberg89}.

The first evolutionary-based approach will be a \textit{GP-based classification} method. This will be useful for two main reasons: first, in order to deal with the data class imbalance \cite{imbalance_techniques_02}, very common in classification problems inside real systems (with real data); and second, to better manage categorical (non-numeric) data, since most of the features/variables and information gathered from the events take these kind of values.

Thus, this approach will be able to manage unbalanced datasets considering a fitness function in which a cost can be associated to the classifier accuracy at every epoch, having a penalty cost when the classifier makes a false negative (an element from the minority class which is classified as belonging to the majority class) \cite{cost_adjustment_07}.
Regarding the type of data, since GP algorithms can manage rule- or tree-based models, it will work perfectly with any categorical variable, yielding a good classifier as it has been made in other works (such as the aforementioned \cite{cost_adjustment_07}).

@@

The second set of EC techniques are, as stated, part of the rule-refinement (or adaptation) process, to be performed by the \textit{Knowledge Compiler} module. These methods will be used for inference and optimisation, and will consider this data as a part of the process:

@@

Another useful information to consider in the refinement/ adaptation process will be:

@@

\item Risk information extracted from the user profile (reputation), e.g. "Did the user received a lot of `denies'/`allows' before?", i.e. "Is he/she trustworthy?". In case the user is not, more restrictive rules would be created for him, otherwise the corresponding rules could be `relaxed' for that user.

@@

\item \textit{GP rule refinement} approach, which will optimise the current set of rules, adjusting the values in the conditions (antecedents), for instance. Thus, some superfluous parts on the rules and even complete rules could be removed or improved, obtaining for instance specialisations or generalisations of existing rules which could mean a better performance.
The evaluation (of the whole set of security rules) will be done considering the number of unlabelled patterns that will be `covered' after the adjustments.

>>>>

The way on how data has been stored and accessed in the companies has completely changed along the years. The use of owned servers being accessed by desktop PCs (and maybe laptops) from inside the company facilities has been transformed into the distribution of these data between a number of machines, even not all belonging to the company, and working over the famous Cloud Computing environment. Besides, being stored in the cloud or not, data are being consulted and modified through a wide amount of devices, some of them owned by the company's users. This is the so-called Bring Your Own Device (BYOD) philosophy. It is becoming highly successful due to the impact that portable devices (such as smartphones and tablets) are having on the market.
Data security and privacy are key factors for a company, thus, to protect them, it is usual to define Organisational Security Policies. Their definition is nowadays a very difficult problem, since the BYOD tendency means that several factors must be considered \cite{Opp_Security11}, most of them previously ignored or non considered in security systems, for instance the current mixture between personal and professional information in these devices (the user could navigate inside social networks where there could be friends and also company partners or clients).

In this scenario several solutions have arisen in order to manage the corporate security. Some solutions are focused only on smartphones, while others are thought for laptops; also some are implemented for a certain platform, but others consider multiplatform. However, most of them try to be non-intrusive (regarding the users' personal data), friendly and easy to use.

Moreover, it has been demonstrated that people are the main hazard regarding the company security \cite{Adams_Users05}, so in this situation, some monitorization and security-aimed applications are arising, aiming to cope with the concept of seamless working experience on different devices. 
This concept is a methodology of work which allows users to start/continue a working session over multiple devices and locations without any significant loss of data. This new situation has a big impact from the point of view of the security \cite{Schu_SecPatterns05}, since the company's data borders have changed in the last years, so now the users can access significant data from outside the enterprise, and possibly through a non absolutely secure channel.

Thus, this work introduces a novel (still in development) system named MUSES, from Multiplatform Usable Endpoint Security, which is a device-independent end-to end user-centric tool, based on a set of security rules defined as specifications of the Corporate Security Policies, but with the ability of `learning' from the user's past behaviour and adapt, even inferring new ones, the set of rules in order to deal with potential future security incidents due to the user's. Then the system will react, in a non-intrusive way, to the potentially dangerous sequence of actions (events) that he or she is conducting at any time.

@@

On the other hand, employee-owned devices like smartphones, can be both used for personal matters at work, and for continuing working outside of the workplace. As the name says, smartphones are more than simple mobile phones, and people who use them in their works have the possibility of maintaining a good balance between work and private life. For this reason, the risk of uncontrolled devices accessing to corporate assets in unsafe conditions, due to the number of new risks which are linked to smartphones \cite{gangula2013survey}, is bigger.

@@

Many tools for companies, as well as for devices, which have adopted the BYOD have been released in the past four years. This way, and
more focused on the enterprise, there are some tools which offer the CSO ways to control the devices which enter into the system, for instance, and also to protect the employees' data by means of data encryption and data protection by requiring the use of strong and secure passwords. Other tools for managing a BYOD situation add to their features guidelines for the CSOs to develop good ISPs. On the other hand, many solutions have been presented which are more focused on the device side, although they implement also the server side. Some of those tools have influenced the development of the MUSES system  itself (Section \ref{sec:muses}), which can complement them, as adds other features that go beyond the state of the art. The present section introduces the most relevant features of these products, as they can be considered related to the MUSES objectives.

@@

One of the first companies who supported the BYOD model was IBM \cite{IBM_tool}, as they recognized the increase of employees who brought their personal smartphones or tablets into the workplace \cite{ibm11}. IBM has different solutions divided into ``technology solutions'', and ``services'', and almost all are are mainly focused on the management of the devices in the system. Then, it might seem clear that the first disadvantage is not including all features in one system, but to make the companies to choose between one or another, or invest even more resources instead.

Having a look into what IBM offers as ``services'', three services can be found:

\begin{itemize} 
	\item \textit{Mobile Enterprise Services}: Is offered as an integrated solution for smartphones, and tablets as well, but then it is divided in different subservices that the company has to acquire separately. Some of them are related to cloud computing, such as events and log management as a cloud security service, cloud backup, or the so-called `hosted vulnerability management'. The last one is specially interesting for the scope of this paper, given the fact that it seems to implement one of the most important advantages of the MUSES system, the self-adaptation. IBM, by performing deeper scans over the security incidents data, either if they were successful (computer forensics analysis) or not (the device was enough secure), they claim to be able to recognise new vulnerabilities - or threats - with enough accuracy. The main difference with MUSES in this case, is that MUSES includes this feature in the system, without the need of purchasing the service separately, and configuring the different parts. Also, the fact that IBM software is not open source as MUSES system is, to verify the success of this feature is highly difficult, as they do not present statistics in their website either.
	\item \textit{Hosted Mobile Device Security Management}: A particular service of the Mobile Enterprise Services. It includes itself a number of subservices, going from tools to make easy to develop secure mobile apps, to an interface for managing the devices in the system, checking their configuration security level. Again, what is offered by IBM as `services', are included as default features in the MUSES system.
	\item \textit{Enterprise Wireless Networks}: Devoted to provide secure channels for communications. This is related to ensuring employees that any connection they might perform to access company assets, is properly secured, at the time it delivers the expected performance. Thus, this service is, in some sense, adaptive to the situation, yet only in terms of a secure connection and not evolving the security policies, like MUSES. 
\end{itemize}

@@

Contrarily to IBM's solution services, Sophos offers a whole product for securing BYOD environments, which is called Mobile Control \cite{Sophos_tool}. It is oriented to IT administration for mobile devices, and it can be deployed in two ways: on-premises, which means that all data and services remain in the servers of the company, and as Software as a Service (SaaS), so that the services are provided through Lightweight Directory Access Control (LDAP). LDAP is an application protocol for accessing and maintaining distributed directory information services over an Internet Protocol (IP) network.

The tool manages all employees' smartphones and tablets from a single-based console. This console monitors the devices since the initial set up and registration, until they log off. Other features are similar to IBM's product, adding some new such as the incorporation of a service called Malware and Web protection, so that there is no need for having a previously installed antivirus in the devices.

With regard to compliance enforcement, the main goal of this Mobile Control tool is to give more importance to company security, instead of giving more flexibility to the users. But it has been demonstrated by Herath and Rao in \cite{SecPolPenalty09} that this practise can result in bad (even not on purpose) users' behaviour. Thus, the company BYOD initiative should include an acceptable use policy to ensure that the users are aware of the measures that the company would take if a device breaches any ISPs. Sophos aims to reach this by doing three main tasks: First, by allowing setting up user and group-based security policies separately. Secondly, by a risk mitigation in which the actions to perform can be set according to the severity of the breach. For minor cases, the company may want to simply inform to the user, but if sensitive data is at risk, a remote wipe may be the chosen option. Finally, compliance check, by constantly performing scans to detect malware and putting the devices in quarantine when they are found infected.

Thus, given the fact that this tool is more focused on the company than in the user, it is worth to compare it to the MUSES system, as the latter tries to focus on the users, being user-centric. This means that is preferred to let the user know about a risky situation, rather than making them feel ``watched''.

@@

The main feature of KNOX is the use of different containers, or environments, for business and personal sides. Each one even includes its own graphical configuration (wallpapers, colours, and so on), in order to be more easily recognised and distinguished by the user.
A password is required in order to enter into the business side and, once \textit{logged in} in this container, no more passwords are needed for the business applications. The applications approved by the company IT department must meet Samsung security standards and, though it has been said that KNOX is a device-focused solution, it offered an API. This API would allow the company to access to over almost 205 predefined IT policies. With regard to information protection methods, data files saved by applications of each environment are encrypted with AES 256-bit algorithm, in such manner that only the appropriate container can access these files. In the same way, the user is not allowed to share data between the two environments.

Features in Blackberry Balance are similar. This security package was announced as a feature of BlackBerry 10. Nevertheless, it is available with BlackBerry Enterprise Service 10, which is a device management, security and app management for BlackBerry, iOS and Android devices. It is necessary to activate BlackBerry Balance for having available some security features, all related or similar to the aforementioned. For instance, a message is displayed when the user tries to copy work data and then paste it into personal apps. Also, a user attempting for actions that are not permitted in the company ISP, or may cause secure work information to be in contact with personal applications, these actions will not be permitted.

@@

This system will be strongly related to the Google Play market, as the applications will be `categorized'. Thus, Android Work will provide the CSOs with a tool to define which apps would be allowed for being installed by the users and considered as corporate-related applications. Several enterprise-security aimed applications will be offered in the market, so the IT would decide which ones will be installed (even as a bundle) in the employees' devices, and also, they can manage the updates of these apps, in order to ensure that all the employees are up to date. For instance, one app could manage the creation of personal/work profiles, so the user just could access corporate assets after login in into the app. Moreover, the enterprise IT department could define specific security policies on these apps, in addition to the inherent protection that the container provides. However, the specifications of this tool \cite{AndroidWork_tool} do not include the ``self-adaptation'' as a feature, as the MUSES system does.

Even though it does not analyse the system information for security rules evolution purposes, it offers policy management. Thus, the system will provide a framework for IT staff to manage business devices, but also, and more importantly, personal devices being used in a BYOD context (The CSO would give the employee an activation code in order to `connect' the smartphone to the enterprise and use this Android Work).
Therefore, IT admins will be able to specify which Google Play apps will be available for users to install through this work profile, including being able to provision apps to specific individuals or groups (inside the company).

@@

\begin{description}
  \item[Mobile Device Management] that is used to manage both user and corporate owned devices, providing support for Android and iOS at the moment. It should be noted that this tool allows the tracking of every enrolled device, as well as obtaining reports and analytics of their use.
  \item[Mobile Application Management] With regard to the software, this platform is able to allow or deny the use of applications on enrolled devices based on the role of the user or policies, thus restricting the use of some apps to certain users.
  \item[Enterprise App Store] This store provides users with both enterprise and public applications approved by the company.
  \item[Mobile Data Security] WSO2 also allows the user's data to be encrypted via password.
\end{description}

@@

\begin{itemize}
\item It offers a central deployment, administration and securing of all mobile devices regardless of their OS. %FERGU: securing como palabra no. security control mejor?
\item It manages and enforces security policies `over the air' for both corporate and employee-owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe/selective wipe corporate data email/contacts on the device if lost or stolen.
\end{itemize}

@@

It is also important to point out that Citrix differentiates between apps used by the user privately and those used for business, locating both on a secure mobile container that is encrypted, and can be locked remotely for safety reasons. Another feature of Citrix is that it uses dedicated micro VPN to connect to Citrix-protected backend services.

Once the main tools and systems in this area have been introduced, in the following section our own system (still in development) is presented, giving an overview on its general architecture. Moreover, we devote a subsection to describe its main features: the use of Data Mining + Machine Learning techniques, and also its self-adaptation ability using Computational Intelligence methods, which compose the real difference with all these tools.

@@

As a summary, this application includes two modules: a \textit{controller} and an \textit{actuator}. The first one monitors the environment (context) and the user's behaviour, translating his/her actions in a sequence of events. These events, along with the patterns defining the user's conduct, are processed by the system in real-time by means of a Risk and Trust Analysis Engine (RT2AE) and an Event Correlation module. Then, a decision is taken in the corporate security operations centre (SOC) side, considering the RT2AE output and the set of security rules adapted to that specific user and context. The corresponding feedback is communicated to the user through the \textit{actuator}, which is also in charge of triggering the recommended actions to stop the user's or application's doings, in case it is required.

The designed MUSES architecture is shown in Figure \ref{fig:architecture}. It is a \textit{client/server} approach in which the \textit{client} program will be installed in every user's mobile or portable device, independently of the platform (operating system and type of device). The \textit{server} side would be installed in the corporate SOC. Both sides are connected through a secure channel (using HTTPS) over the Internet. In that figure just the high-level components in each part are shown, along with the information flows labelled as 'Info XX'. %FERGU: esto no se si es de muses y posiblemente esté por terminar, pero quitad las comillas raras

@@

The rationale for this approach is based on two main reasons: 1) there is a need for a high computational power in order to perform the event correlation and self-adaptation processes, so a quite powerful machine should be used (server); 2) there are two clearly separated parts in the system, namely the users (client) and the enterprise (server).

The system considers two working modes for every device: online and offline. In the \textit{online} mode the device can connect with the MUSES server, so it can request the server to make a decision. On the other hand, in \textit{offline} mode the server cannot be reached by the device (there is not an available connection between them), so all the decisions should be made locally.
This will be done using an up-to-date copy on the Decision Table inside the Local database, which will contain decisions for any type of detected event, along with the default actions to be performed if no rule is met. Anyway, in this mode, the gathered information by the sensors in the device will be stored for later submission to the server side (when a connection is available), in order to be processed in the knowledge refinement process.
Moreover, when switching from offline to online mode, the device will receive an updated copy of the Decision Rules (the so-called Device Policies), as a kind of antivirus database, in order to work again perfectly in offline mode (if the connection is lost).

All data in MUSES will be encrypted. %FERGU: esta frase esta un poco forever alone aquí, falta algo por mencionar? Que tecnología se usará, cómo se encriptará y todo eso.

@@

The rest of the components are: the \textit{MUSES User Interface (UI)}, the application through which the user interacts with the system; and the \textit{Connection Manager} which controls the communications between client and server sides. 
In addition, there are two types of applications considered in this system: on the one hand the \textit{MUSES Aware App}, which is an application adapted to MUSES, so the system can directly interact with it (monitoring and acting). This application must be implemented using the MUSES API (Application Program Interface)\footnote{The MUSES API will be defined in the project, so for every application desired to be MUSES Aware, it should be implemented using this.}. On the other hand the \textit{Non MUSES Aware App} is that which MUSES cannot directly interact with. Usually it will be accessed through the operating system (OS). So the type of information to gather from them and also the type of actions to do over them will vary depending on the OS. In most of the cases MUSES could just detect/recognize very specific events, and also do very limited actions to prevent or block them (maybe showing a screen on the top until something is accomplished by the user). This is due to the usual forbiddance or limitation of interactions with other applications that the OSs have. %FERGU: forbiddance parece que existe, pero el corrector me dice que no. Prohibition mejor?

@@

There are some other components, namely: the \textit{Security Policies/Risk Management} tool, that lets the company's chief security officer (CSO) to define and manage Security Policies and Rules in a friendly way. It also lets the management of risk-related information, useful for the RT2AE process, such as the assets' values. The \textit{Privacy Enhancing System} which is a module aimed to fit with the legal compliance of the system regarding the user's data anonymisation and retaining of these data in the system. The \textit{User, Context, Event Data Receiver} is devoted to receive data from the device side (events, user-related, etc.) and to distribute them among the components (storing in the database, and requesting the Event Processor to start the correlation task). Finally, there is another \textit{Connection Manager} which controls the communication with the device side. %FERGU: no usar etc en artículos, mejor poner: (events, user-related LOQUESEA, among others)

@@

Another important fact is that MUSES will count with a human controller, normally the Chief Security Officer (CSO) of the company, who will supervise the system activity by means of logs or feedback messages. Thus, adapted and inferred security rules will not be, in principle, directly added to the current set of rules. Instead, they will be proposed as draft rules to this controller, in order that he/she will accept them if they are interesting and correct. It is planned that the system will be able to `learn' from this decisions, so after a so-called training or `warm up' period, the rules would be directly accepted (included in the set) or rejected autonomously by the MusKRS.

@@

The DM/ML techniques will process the so-called patterns, which in this context correspond to events (and their related information) produced by the users' interactions with the system. As stated, these events will be captured from the sensors in the users' devices and transmitted to the server to be processed and stored in the system database. The methods to be applied are:

@@

\item \textit{Classification} \cite{classification_67}: 
This technique tries to train a model (classifier) able to associate every pattern in the dataset to a class, so that the model could be used for assigning a class for further incoming patterns with an unknown category.
For instance, it could look for events (patterns) that had been previously  marked as `allowed' or `denied' (according to the ISPs). When a new event arises, if it has not an assigned decision, the classifier should provide one based on the similarity with previous (and already labelled) patterns.

@@

\item \textit{Feature Selection} \cite{FeatureSelection_Guyon03}: 
It consists on extract the most important features/variables from the data. This could be useful if we want to discard non-key features, which could be interesting in order to reduce the database weight, for improving the performance of other techniques (such as classification or clustering), and even to improve the performance of the whole system, since less information would be gathered and transmitted.

@@

MUSES will use different EC approaches, initially, two in the DM/ML part of the process, and three in the rule-refinement/adaptation phase. They are based on Genetic Programming (GP) \cite{GP_Koza92}, and Genetic Algorithms (GAs) \cite{GAs_Goldberg89}.

The first evolutionary-based approach will be a \textit{GP-based classification} method. This will be useful for two main reasons: first, in order to deal with the data class imbalance \cite{imbalance_techniques_02}, very common in classification problems in real systems (with real data); and second, to better manage categorical (non-numeric) data, since most of the features/variables and information gathered from the events take these kind of values.

Thus, this approach will be able to manage unbalanced datasets considering a fitness function in which a cost can be associated with the classifier accuracy at every epoch, having a penalty cost when the classifier makes a false negative (an element from the minority class which is classified as belonging to the majority class) \cite{cost_adjustment_07}.
Regarding the type of data, since GP algorithms can manage rule- or tree-based models, it will work perfectly with any categorical variable, yielding a good classifier as it has been made in other works (such as the aforementioned \cite{cost_adjustment_07}).

@@

The second set of EC techniques is, as stated, part of the rule-refinement (or adaptation) process, to be performed by the \textit{Knowledge Compiler} module. These methods will be used for inference and optimisation, and will consider this data as a part of the process:

@@

Other useful information to consider in the refinement/ adaptation process will be:

@@

\item Risk information extracted from the user profile (reputation), e.g. ``Did the user received a lot of `denies'/`allows' before?'', i.e. ``Is he/she trustworthy?''. In case the user is not, more restrictive rules would be created for him, otherwise the corresponding rules could be `relaxed' for that user.

@@

\item \textit{GP rule refinement} approach, which will optimise the current set of rules, adjusting the values in the conditions (antecedents), for instance. Thus, some superfluous parts of the rules and even complete rules could be removed or improved, obtaining for instance specialisations or generalisations of existing rules which could mean a better performance.
The evaluation (of the whole set of security rules) will be done considering the number of unlabelled patterns that will be `covered' after the adjustments.
